---
title: "DM covariates, lags"
author: "Glenda Yenni"
output: github_document
---
  
  Patterns in DM abundance related to precipitation, temperature, and ndvi


```{r global options, include = FALSE}
knitr::opts_chunk$set(echo=FALSE, include = TRUE, warning=FALSE, message=FALSE)

library(dplyr)
library(ggplot2)
source("../R/plotting.R")
```

```{r get data}
ndvi_dat <- portalr::ndvi(level = "monthly", fill = FALSE) %>%
  mutate(year = lubridate::year(date), month = lubridate::month(date)) 

weath_dat <- portalr::weather(level = "monthly", fill = FALSE) %>%
  select(year, month, mintemp, maxtemp, meantemp, precipitation) %>%
  left_join(ndvi_dat, by = c("year", "month"))

rodent_dat <- portalr::abundance(time = "date") %>%
  select(censusdate, DM) %>%
  mutate(censusdate = lubridate::ymd(censusdate)) %>%
  mutate(year = lubridate::year(censusdate), month = lubridate::month(censusdate)) %>%
  left_join(weath_dat, by = c("year", "month"))

```

## The time series
Monthly, because the ndvi we have is monthly
```{r time series}

theme_set(theme_minimal())
p1 <- ggplot(data = ndvi_dat, aes(x = date)) + 
  geom_line(aes(y = meantemp, color = "meantemp")) +
  geom_line(aes(y = mintemp, color = "mintemp")) +
  geom_line(aes(y = maxtemp, color = "maxtemp")) +
  scale_color_manual(name="",
                     values = c("maxtemp"="red","meantemp"="pink","mintemp"="blue")) +
  labs(x = '', y = 'Air Temp (C)') +
  theme(legend.position = "top") 

p2 <- ggplot(data = ndvi_dat, aes(x = date)) + 
  geom_line( aes(y = ndvi), color = "green") +
  xlab('')

p3 <- ggplot(data = ndvi_dat, aes(x = date)) + 
  geom_line(aes(y = precipitation), color = "lightblue") +
  xlab('date') + ylab('Precip (mm)')

cowplot::plot_grid(p1,p2,p3,nrow=3)

```

## NDVI and Temperature
```{r ndvi vs temp}

plot_temp_ndvi(ndvi_dat, lag = 0, method = "loess") 
plot_temp_ndvi(ndvi_dat, lag = 1, method = "loess")
plot_temp_ndvi(ndvi_dat, lag = 3, method = "loess")
plot_temp_ndvi(ndvi_dat, lag = 6, method = "loess")
plot_temp_ndvi(ndvi_dat, lag = 12, method = "loess")
```

## NDVI and Precipitation
```{r ndvi vs precip}

plot_precip_ndvi(ndvi_dat, lag = 0, method = "loess")
plot_precip_ndvi(ndvi_dat, lag = 1, method = "loess")
plot_precip_ndvi(ndvi_dat, lag = 3, method = "loess")
plot_precip_ndvi(ndvi_dat, lag = 6, method = "loess")
plot_precip_ndvi(ndvi_dat, lag = 12, method = "loess")

```

## NDVI, Mean Temperature, and Precipitation
```{r ndvi vs temp and precip}

plot_glm_ndvi(ndvi_dat, lag = 0)
plot_glm_ndvi(ndvi_dat, lag = 1)
plot_glm_ndvi(ndvi_dat, lag = 3)
plot_glm_ndvi(ndvi_dat, lag = 6)
plot_glm_ndvi(ndvi_dat, lag = 12)

```

## Forecast NDVI and Temperature
Currently we only have NDVI through 2013. Below are comparisons between the NDVI prediction via autocorrelation (w/ seasonality) and the predictions from the best model above (`ndvi~lag(meantemp,3)*lag(precipitation,3)`) (darkgreen)
```{r forcast ndvi vs temp and precip}

fit=glm(ndvi~lag(meantemp,3)*lag(precipitation,3),data=ndvi_dat)
preds = predict(fit,newdata = ndvi_dat_filled[391:459,])

ndvi_dat_preds <- ndvi_dat_filled %>%
  mutate(pred = c(rep(NA, 390), preds))

theme_set(theme_minimal())
ggplot(data = ndvi_dat_preds, aes(x = date)) + 
  geom_line(aes(y = ndvi), color = "green") +
  geom_line(aes(y = pred), color = "darkgreen") 

ggplot(data = ndvi_dat_preds) + 
  geom_point(aes(x = ndvi, y = pred), color = "green") +
  geom_abline(intercept=0, slope=1) +
  labs(x = "autocorrelation predicted", y = "temp*ppt predicted")

```